---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
# Welcome to hihai explorer

<!-- badges: end -->

## Overview

The `hihai` package transforms heathcare data into an interactive tools. It provides two tidy datasets and Shiny app to explore burden of Healthcare-Associated Infections (HAIs) in Germany compared to other EU/EEA countries.

## Data

The data is based on BHAI package that was collected from **2011 ECDC Point Prevalence Survey** covering 46 German hospitals (9,626 patients) and 29 EU/EEA countries (273,753 patients).

**Hihai** makes it easy to:

- Explore HAI burden across five infection types

  • **HAP** - Health-Associated Pneumonia
  
  • **SSI** - Surgical Site Infection
  
  • **BSI** - Bloodstream Infection
  
  • **UTI** - Urinary Tract Infection
  
  • **CDI** - Clostridioides difficile Infection

- Analyse frequency vs severity to identify high-impact infections
- Compare Germany's performance against EU/EEA countries
- Visualise trends with interactive charts

## Installation

You can install the development version from:
``` r
# install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-arisarath")
```

### Launch Shiny app
```{r app, eval=FALSE}
hihai::launch_app()
```

The Shiny app features:

- **Getting started** : Introduction and field explanations
- **Explore** : Bar chart comparing HAIs by different metrics
- **Severity** : Bubble chart showing severity
- **Comparison** : Germany vs EU/EEA clustered chart

## Learn more
See the [examples][examples] page for full tutorial.

[examples]: https://ETC5523-2025.github.io/assignment-4-packages-and-shiny-apps-arisarath/articles/hihai_explorer.html

## Source
- Zacher et al. (2019). *Application of a new methodology and R package reveals a high burden of healthcare-associated infections (HAI) in Germany compared to the average in the European Union/European Economic Area, 2011–2012.* Eurosurveillance, 24(46). https://doi.org/10.2807/1560-7917.ES.2019.24.46.1900135

## License
MIT License.

## Contributing

Please [submit an issue][issues] to report bugs or ask questions.

[issues]: https://github.com/ETC5523-2025/assignment-4-packages-and-shiny-apps-arisarath/issues
