---
title: "hihai explorer"
author: "Arisara Therdthianwong"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hihai explorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hihai)
```

## Introduction

Imagine going to the hospital for treatment but ended up leaving with another infection.

Healthcare-associated infections (HAIs) are a major health burden in Germany, affecting approximately 1 in 25 hospitalised patients. These infections cause significant illness, deaths, and health costs every year.

The `hihai` package provides clean datasets and interactive Shiny app to explore, visualise, and interpret the burden of HAIs in Germany using the data from **2011 ECDC Point Prevalence Survey**.

## About the data

The package uses two cleaned datasets:
1. `hai_data`: German PPS (46 German hospitals with 9,626 patients) with 95% uncertainty interval (UI).
2. `hai_comparison`: Germany vs EU/EEA rates using ECDC PPS (1,149 hospitals with 273,753 patients) with 95% uncertainty interval (UI).

## What's measured

- hai_type: abbreviation of 5 HAI types
- hai_name: name of 5 HAI types
- cases: Annual number of infections
- deaths: Attributable mortality
- dalys: Disability-adjusted life years (YLL + YLD)
- yll: Years of life lost
- yld: Years lived with disability

*All estimates include lower and upper bound for 95% UI*

### Exploring Germany's HAI burden:
```{r}
head(hai_data)
```

You can also explore international comparison:
```{r}
show(hai_comparison)
```

## Let's explore the data with Shiny App

This application contains 3 main tabs; explore, severity, and comparison.
- Explore view: Examine HAI frequency with customisable metrics
- Severity view: Visualise the relationship between frequency and impact
- Comparison view: Contrast German rates with EU/EEA averages

Launching the App:
```{r launch-app, eval=FALSE}
launch_app()
```

## Tab1: Explore - Understanding frequency

**Question answered:** How common each HAI type is and what impact do they have?

### What you see
This tab focuses solely on Germany (German PPS data).

The **Bar chart** shows the frequency of HAIs in selected metric (cases, deaths, or DALYs) for each HAI type in descending order.

### Interactive controls
1. Metric dropdown: 
- Cases: How many infections occurred
- Deaths: How many people died from HAIs
- DALYS: Overall health burden
- YLLs: Years of life lost
- YLDs: Years lived with disability

2. Check box
- Check/uncheck to filter specific infections
- Useful for comparing side-by-side data

3. 95% UI toggle
- Shows error bars for statistical uncertainty
- Lower and upper bound where the true value likely falls

### How to read the chart
- X-axis: HAI types (UTI, HAP, SSI, CDI, BSI)
- Y-axis: Changes based on your selected metric
- Bar height: Value for that metric
- Error bars: 95% Uncertainty interval
- Colors: Each HAI type has a distinct color

### Data table
Table shows the actual number of cases, deaths, DALYs, YLLs, and YLDs.
You can sort by clicking column headers (Descending or Ascending).

### Key findings
At the bottom, the app automatically generates a **ranking summary**.

*Example: When we look at cases: total = 476,936 cases. 1st rank is Urinary Tract Infection (UTI) with 214,150 cases, accounting for 44.9% of total infections.*

This summary updates based on your selected metric, making it easy to capture findings. 

**What you learn:** UTI is the most common type of HAI, but HAP causes the most harm overall.

## Tab2: Severity - Visualising the impact

**Question answered:** Which infections are most deadly relative to how common they are?

Frequency alone doesn't tell the full story. To understand impact, the package combine cases, deaths, and DALYs using a bubble chart. 

### What you see

The **Bubble chart** reveals the relationship between:
- Frequency (how common)
- Mortality (how deadly)
- Burden (overall impact)

### How to read the bubble chart
- X-axis: Annual number of cases (Further right => more common)
- Y-axis: Annual attributable deaths (Higher => more deadly)
- Bubble size: DALYs (bigger bubble = higher overall burden)
- Color: Each HAI type has a distinct color
- Label: Abbreviation of HAI type (UTI, HAP, SSI, CDI, BSI)

### Interactive features
**Hover over any bubble** to see a detailed statistics:

*Example tooltip:* 
```
Urinary Tract Infection
Cases:214,150 (175,086 - 253,524)
Deaths: 3,664 (1,462 - 7,533)
DALYs: 66,701 (27,890 - 128,543)
```

### Interpreting the chart
**Upper Right (e.g., UTI)**
- Many cases (far right)
- Moderate deaths (middle height)
- Large bubble (high total burden due to volume)

**Upper left (e.g., BSI HAP)**
- Fewer cases (left side)
- High deaths (high up)
- Large bubble (high burden despite lower frequency)

### Key findings
Frequency alone is misleading, BSI has only 26,976 cases but 3,905 deaths while UTI has 214,150 cases but 3,664 deaths.

**What you learn:** Despite having fewer cases, BSI causes similar total deaths.

## Tab3: Comparison - Germany VS EU/EEA

**Question answered:** How Germany's HAI burden perform comparing to other European countries?

### What you see
A **side-by-side bar chart** comparing:
- Red bars: Germany (German PPS)
- Blue bars: EU/EEA Average (ECDC PPS)

### Interactive controls
1. Metric dropdown:
- Cases: How many infections occurred
- Deaths: How many people died from HAIs
- DALYS: Overall health burden

2. Population scale slider:
- Adjust from 1,000 to 100,000
- Changes the "per X population" calculation
- Default: 100,000 population

### How to read the chart
- X-axis: HAI types (UTI, HAP, SSI, CDI, BSI)
- Y-axis: Rates per 100,000 population (adjust according to slider)
- Error bars: 95% Uncertainty interval

### Key findings
Below the chart, the app automatically generates a **automatic interpretation** that compares each HAI type based on metric you choose.

```
Example: When we look at (100,000 population), 
Comparing German PPS to EU/EEA rates for DALYs:
- CDI: Germany 159% higher (25.9 vs 10)
- UTI: Germany 45% higher (82.6 vs 57.1)
- HAP: Germany 22% lower (86.1 vs 109.8)
- BSI: similar rates (Germany: 72.2 vs EU/EEA: 76.2)
- SSI: similar rates (Germany: 35.7 vs EU/EEA: 35.1)*
```

**What you learn:** German shows higher burden across most of the metrics, despite having good hospital infetion protocol. Reasons include high hospitalisation rates and large bed capacity, which create more exposure opportunities.

---

## FAQ & Common Questions
### Q1: Why sometimes the number is not exactly match?
**Ans:** All row is calculated independently using Monte Carlo simulations. Due to rounding in UI, it may result in differ from sum of individual HAI types.

### Q2: What does 95% UI means?
**Ans:** It is an uncertainty interval where confidently 95% true value lies within this range.

### Q3: Which metric should I use?
**Ans:** 
| Question | Use |
|----------|-----|
| "How common is this?" | Cases |
| "How deadly is this?" | Deaths |
| "What's the overall burden?" | DALYs |
| "How many years of life are lost?" | YLLs |
| "How much disability does this cause?" | YLDs |

---

## Summary: key takeaways
### 1. The problem is significant
HAI are major health burden in Germany that causes more than **478,000 cases**, over **16,000 deaths** and **249,000 DALYs** each year.

### 2. Frequency VS Severity
UTI is the most common type of HAI contributing to 45% of total cases but causes moderate severity to human. Meanwhile HAP and BSI are less common with roughly 22% and 6% of overall frequency but causes 51% of total mortality.

### 3. Germany VS Europe
Germany has a higher population burden than EU/EEA average. Due to high hospitalisation rates and bed capacity accelerate exposure overall.

### 4. Suggestions
According to WHO, 20-30% of HAIs are preventable through better hygiene, infection control, and shorter hospital stays. Prevention strategies should prioritise infections that cause the greatest harm, not just the most frequent one.

## Learn more

### Access package data
```{r help, eval=FALSE}
# Load datasets
data("hai_data")
data("hai_comparison")

# View documentation
?hai_data
?hai_comparison
?launch_app
```

### Additional resources

- Package help: Type `?hihai` in R console
- Original study: Zacher et al. (2019) (https://doi.org/10.2807/1560-7917.ES.2019.24.46.1900135)
 

## References

Zacher, B., Haller, S., Willrich, N., Walter, J., Sin, M. A., Cassini, A., Plachouras, D., Suetens, C., Behnke, M., Gastmeier, P., Wieler, L. H., & Eckmanns, T. (2019). Application of a new methodology and R package reveals a high burden of healthcare-associated infections (HAI) in Germany compared to the average in the European Union/European Economic Area, 2011 to 2012. Eurosurveillance, 24(46).
https://doi.org/10.2807/1560-7917.es.2019.24.46.1900135
